---
title: "PsyArXiv"
author: "Jihan Ryu"
date: "2023-10-17"
output: html_document
---

# 1. Load packages

```{r}
knitr::opts_chunk$set(echo = FALSE)
install.packages(c("ggplot2","dplyr","tidyr","ggcorrplot","ggpubr","statsr","lme4","lmerTest","magick"))
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggcorrplot)
library(ggpubr)
library(statsr)
library(lme4)
library(lmerTest)
library(magick)
```

# 2. Upload datasets in correct format and descriptive stats

```{r}
# Load datasets

data022822 <- read.csv("TG_Raw_76.csv")

# Factor data types into correct formet

factors<-c("Role","Dyad","Modality","Diagnosis","ID","Sex","Inherited","T_Experience","Medication")
integers<-c("Perception_Mismatch","Total_Alliance","Visits","Age","Task","Goal","Bond","Indiv_Total","Intrusiveness","Previous_Alliance","Lifetime_Therapist_Number","Closeness","Dependence","Anxiety","Avoidance","RFQc","RFQu","Investor_Total","Trustee_Total","Scores_Total","rt1","rt2","rt3","rt4","rt5","rt6","rt7","rt8","rt9","rt10")
numerics<-c("r3","r4","r5","r6","r7","r8","r9","r10","d_rf3","d_rf4","d_rf5","d_rf6","d_rf7","d_rf8","d_rf9","d_rf10","rf1","rf2","rf3","rf4","rf5","rf6","rf7","rf8","rf9","rf10")
data022822[,factors]<-lapply(data022822[,factors], factor)
data022822[,integers]<-lapply(data022822[,integers], as.integer)
data022822[,numerics]<-lapply(data022822[,numerics], as.numeric)
data022822$T_Experience <- factor(data022822$T_Experience, levels=c("<10","11+","100+"))

# Change missing, N/A, or #VALUE! data values into NA

data022822[data022822 == "N/A" | data022822 == "#VALUE!" | data022822 == ""] <- NA

# Create correlation matix among numerically valued variables

temp_corr<-data022822[c(38:49,51:58,60:74),] #Subject 26 and 44 (patients, i.e. row 50 and 59) data are missing
temp_corr<-temp_corr %>% 
  select_if(is.numeric)
temp_corr<-temp_corr[,1:15]
temp_corr$Indiv_Total_Counterpart = temp_corr$Total_Alliance - temp_corr$Indiv_Total # mutate a new variable called Indiv_Total_Counterpart, i.e. therapeutic alliance rated by their therapy partner (patient or therapist) in the pair

corr.vars <- cor(temp_corr, method="pearson")
p.mat <- cor_pmat(temp_corr)

# Create Supplementary Fig. 1

corr_plot<-ggcorrplot(corr.vars, 
           hc.order = TRUE,
           type = "lower",
           p.mat = p.mat)

ggsave("corr_plot.png")

```

# 3. Mutate and extract average repayment fraction and reaction time parameters from Trust Game

```{r}
# Mutate and generate mean reapyment values

data022822 <- data022822 %>% 
  rowwise() %>% 
  mutate(rf_avg = mean(c(rf1, rf2, rf3, rf4, rf5, rf6, rf7, rf8, rf9, rf10), na.rm = TRUE),
         rt_avg = mean(c(rt1, rt2, rt3, rt4, rt5, rt6, rt7, rt8, rt9, rt10), na.rm = TRUE))

# Summarize mean repayment by roles

rf_avg_p_mean<-colMeans(data022822[data022822$Role=="P" & is.nan(data022822$rf_avg)==FALSE,][,'rf_avg'])
rf_avg_t_mean<-colMeans(data022822[data022822$Role=="T",][,'rf_avg']) # this value should be 0.571

# For Supplementary Table 2: Create a separate dataframe for social reprocity information received by participants (note: reciprocity value for r1 and r2 cannot be calculated as it requires player beavhior in the previous two rounds)

investment<-data022822[,c("Role","ID","r3","r4","r5","r6","r7","r8","r9","r10","I1","I2","I3","I4","I5","I6","I7","I8","I9","I10")]

# Count the number of positive/negative reciprocity rounds received by looping through values in r3-r10 

# Set counter for the number of positive/negative reciprocity rounds to zero

investment$pos = 0
investment$neg = 0

# Loop

for (j in 1:nrow(investment)){
    for (k in 3:10){
      if (is.na(investment[j,k])) {investment[j,k]=FALSE} #make sure all arguments are operating under TRUE/FALSE able environment
       else{
        if (investment[j,k]>0){
        investment[j,]$pos = investment[j,]$pos + 1
        }
        if (investment[j,k]<0){
        investment[j,]$neg = investment[j,]$neg + 1
        }
      }   
    }
}

# Mean of raw investment received by participant throughout the 10-round

investment<-investment %>% 
  rowwise() %>% 
  mutate(i_avg = mean(c(I1, I2, I3, I4, I5, I6, I7, I8, I9, I10), na.rm=TRUE))

# Split dataframes into patients and therapists

data022822_p<-data022822[data022822$Role=="P",]
data022822_t<-data022822[data022822$Role=="T",]

# Template correlation/inference test across clinical variables and social reciprocity differences between roles

cor.test(data022822_p$rf_avg, data022822_p$Closeness, method=c("pearson")) 
cor.test(data022822_t$rf_avg, data022822_t$Indiv_Total, alternative="two.sided", method=c("pearson")) 
inference(y=neg, x=Role, data=investment, statistic="mean", type="ci", conf_level=0.95, method="theoretical", alternative="two.sided")

# Create Figure 1

# Fig 1a: Trust Game cartoon
tg <- image_read('~/Data/Trust Game Data/TG_exp.png') 
tg <-image_annotate(tg, "A", size=100, gravity = "northwest", color = "black")
tg <- image_scale(tg, "800")
print(tg)

# Fig 1b: Mean pepayment frequency distribution between therapists and patients

Fig1b <- ggplot(data022822)+
  geom_histogram(data=data022822[data022822$Role=="P",],
                 aes(x=rf_avg), fill="Orange", alpha = 1, position="identity")+
  geom_vline(aes(xintercept = rf_avg_p_mean), colour="Orange", linetype=5)+
  geom_histogram(data=data022822[data022822$Role=="T",],
                 aes(x=rf_avg), fill="Blue", alpha = 0.8, position="identity")+
  geom_vline(aes(xintercept = rf_avg_t_mean), colour="Blue", linetype=5)+
  labs(y="Frequency", x="Mean Repayment",
       tag="B")+
  scale_fill_manual(name="Role", 
                    values=c("Orange", "Blue"), 
                    labels=c("Patient", "Therapist"))

# Fig 1c: Correlation between mean repayment frequency and therapeutic alliance

Fig1c <- ggplot(data022822, aes(rf_avg, Indiv_Total, color=Role))  +
  geom_point() +
  geom_smooth(method=lm)+
  scale_color_manual(name="Role",
                     values=c("Orange", "Blue"), 
                     labels=c("Patient", "Therapist")) +
  labs(y="Alliance", x="Mean Repayment",
       tag="C")

# Combine Fig 1a, b, c into Fig 1.

Fig1bc <- ggarrange(Fig1b, Fig1c, ncol=2, nrow=1, common.legend = TRUE)
ggsave("Fig1bc.png", width=4, height=2)
tg_2 <-image_read('~/Data/Trust Game Data/Fig1bc.png')

tg_full <-c(tg, tg_2)
Fig1 <-image_append(image_scale(tg_full, "700"), stack=TRUE)
image_write(Fig1, path = "Fig1.png", format = "png")

# Supplementary Figure 2: Patients' average repayment fraction is not correlated with attachment closeness subscale.

FigS2a <-ggplot(data022822_p, aes(rf_avg, Closeness))  +
  geom_point() +
  geom_smooth(method=lm)+
  scale_color_manual(values=c("Orange"), 
                     labels=c("Patient")) +
  labs(y="Attachment Closeness", x="Mean Repayment")

```

# 4. Prepare dataframes for reciporicty, changes in repayment, repayment fraction per individual for statistical analysis

```{r}
# First pivot, r3-10, drf3-10, and rt3-10 per ID along the rows

r<-data022822 %>% 
  pivot_longer(
    cols=c("r3", "r4", "r5", "r6", "r7", "r8", "r9", "r10"),
    names_to="round",
    values_to="r",
    values_drop_na=FALSE
  )
d_rf<-data022822 %>% 
  pivot_longer(
    cols=c("d_rf3", "d_rf4", "d_rf5", "d_rf6", "d_rf7", "d_rf8", "d_rf9", "d_rf10"),
    names_to="round",
    values_to="d_rf",
    values_drop_na=FALSE
  )
rf<-data022822 %>% 
  pivot_longer(
    cols=c("rf3", "rf4", "rf5", "rf6", "rf7", "rf8", "rf9", "rf10"),
    names_to="round",
    values_to="rf",
    values_drop_na=FALSE
  )

# Combine pivoted dataframes for reciprocity values, changes in repayment fraction in that round, repayment fraction into a new data frame, i.e. 'mixed' horizontally.

mixed<-cbind(r[,c(1:26,45:47,69:72)],d_rf[,72],rf[,72]) #see if all columns you are interested in analyzing are in place before combining.

# Check distribution of all parameters visually before running further stats

ggplot(mixed, aes(x=r, color=ID))+
  geom_histogram(binwidth=0.1)
ggplot(mixed, aes(x=d_rf, color=ID))+
  geom_histogram(binwidth=0.1)
ggplot(mixed, aes(x=rf, color=ID))+
  geom_histogram(binwidth=0.05)
```

# 5. Compare mean fractional change in repayment between benevolent rounds (reciprocity \>0) and malevolent rounds (reciprocity \<0)

```{r}

# Binarize reciprocity into malevolence vs benevolence

mixed$intention<-ifelse(mixed$r>0,"Benevolence",
                          ifelse(mixed$r<0, "Malevolence",NA))

# Delete round observations with NA in either r or d_rf column 

mixed_filtered <- mixed %>% 
  filter(!is.na(r) & !is.na(d_rf)) #75 rounds should be lost due to NA data. Checkpoint.

mixed_filtered$intention<-as.factor(mixed_filtered$intention)
summary(mixed_filtered$intention) #26 neutral intentions. 

# Fig 2: Plot changes in repayment fraction ~ social reciprocity

d_rf_intention<-mixed[is.na(mixed$intention)==FALSE,] %>% 
  group_by(intention, Role) %>% 
  summarise(sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n())%>%
  mutate(se=sd/sqrt(n))

Fig2 <- ggplot(d_rf_intention, aes(x=intention, y=mean_d_rf, fill=Role))+
  geom_bar(stat="identity", width=0.2, position=position_dodge(width=0.2)) +
  geom_errorbar(
    aes(x=intention, ymin=mean_d_rf-se, ymax=mean_d_rf+se, width=0.04), position=position_dodge(width=0.2))+
  labs(x="Therapy partner's intention", y="Mean Fractional Change in Repayment")+
  scale_fill_manual(name="Role", 
                    values=c("Orange", "Blue"), 
                    labels=c("Patient", "Therapist"))

ggsave("Fig2.png", width=5, height=3) 

# Divide into different dataframes based on their roles (patients vs therapists)

mixed_t <- mixed_filtered %>% 
  filter(Role == "T")
mixed_p <- mixed_filtered %>% 
  filter(Role == "P")

# Before testing significant differences between binary conditions, remove neutral rounds 

mixed_p_without_neutral <-mixed_p[is.na(mixed_p$intention)==FALSE,]
mixed_t_without_neutral <-mixed_t[is.na(mixed_t$intention)==FALSE,]

mixed_p_without_neutral %>% 
  filter(intention=="Benevolence")->tbv
mixed_p_without_neutral %>% 
  filter(intention!="Benevolence")->tml

var.test(tbv$d_rf, tml$d_rf) #p>0.05 in variance test indicates homoscedasticity 

# Run t.test between two conditions within the same role

t.test(d_rf~intention, paired=FALSE, vars.equal=TRUE,
       data=mixed_p_without_neutral) # Note: not exactly matched or paired, i.e. some patients had more malevolent rounds than benevolent rounds, error in complete.cases(x,y): not all arguments have the same length. only patient data show sig differences between conditions

# Run t.test between two roles within same condition

mixed_filtered[mixed_filtered$intention=="Benevolence",] %>% 
  filter(Role=="T")->tT
mixed_filtered[mixed_filtered$intention=="Benevolence",] %>% 
  filter(Role!="T")->tP

var.test(tT$d_rf, tP$d_rf) #p>0.05 homoscedacticity for benevolence, but p=0.03 for malevolence

t.test(d_rf~Role, paired=FALSE, vars.equal=TRUE,
       data=mixed_filtered[mixed_filtered$intention=="Benevolence",]) 

```

# 6. Test the effect of therapeutic alliance on repyament fraction and changes in repayment fraction for patients

```{r}

# Identify median alliance in patient group to binarize alliance variables into high vs. low alliance

test_mixed_p<-mixed_p %>% 
  distinct(ID, .keep_all = TRUE) # trim down the dataframe to isolate clinical variable that is unique to each partient

summary(test_mixed_p$Indiv_Total) #median should be 74. Check point. 

# NOTE: we will work with neutral rounds as well, now that we are dealing with continuous variables.
# Binarize alliance, attachment avoidance, attachment anxiety scores into "Top Half" or "Top Bottom" group

mixed_p$alliance_Q<-ifelse(mixed_p$Indiv_Total>74,"Top_Half",
                               "Bottom_Half")
mixed_p$avoidance_Q<-ifelse(mixed_p$Avoidance>28,"Top_Half",
                               "Bottom_Half")
mixed_p$anxiety_Q<-ifelse(mixed_p$Anxiety>22,"Top_Half",
                               "Bottom_Half")
 
# Make binarized clinical variables into factor data type

factors_2 <- c("alliance_Q", "avoidance_Q", "anxiety_Q")
mixed_p[,factors_2]<-lapply(mixed_p[,factors_2], factor)

# Run 2 (high vs. low alliance) x 2 (benevolence vs. malevolence) ANOVA test on changes in repayment fraction for patients

aov_d_rf_alliance <- aov(d_rf ~ intention * alliance_Q, data=mixed_p[is.na(mixed_p$intention)==FALSE,])
summary(aov_d_rf_alliance) # not significant

# Summarize the mean changes in repayment fraction across social reciprocity and alliance conditions 

d_rf_by_alliance_Q <-mixed_p[is.na(mixed_p$intention)==FALSE,] %>% 
  group_by(alliance_Q, intention) %>% 
  summarise(sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n())%>%
  mutate(se=sd/sqrt(n))

# Run correlation test to examine relationship between in mean fractional change in repayment and individual alliance rating within each social reciprocity condition

d_rf_by_intention_p <-mixed_p[is.na(mixed_p$intention)==FALSE,] %>%
  group_by(intention, ID) %>% 
  summarise(ID=ID, sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n(), Indiv_Total=Indiv_Total)%>%
  mutate(se=sd/sqrt(n)) %>% 
  distinct(ID, .keep_all = TRUE)

attach(d_rf_by_intention_p[d_rf_by_intention_p$intention=="Benevolence",])
cor.test(mean_d_rf, Indiv_Total, alternative="two.sided", method="pearson") # not significant for linear relationship in either benevolent or malevolent condition

```

# 7. Build linear mixed effect models of predicting changes in repayment fraction in patients as a function of social reprocity (as continuous variable, not categorical) and clinical variables (e.g. attachment style)

```{r}

# Remove outliers of extreme malevolence or extreme benevolence, defined by r value cut off -5 and 5, and create a separate data frame for LMM modeling.

mixed_p_lme4<-mixed_p[mixed_p$r > -5 & mixed_p$r < 5,] 
str(mixed_p_lme4) #237 observation points. Checkpoint.

# Create two columns for social reciprocity variables (i.e. r_high_alliance, r_low_alliance) to store as separate predictors depending on which alliance group an individual patient belongs to

mixed_p_lme4$r_high_alliance <- ifelse(mixed_p_lme4$alliance_Q=="Top_Half",
                                  mixed_p_lme4$r, 0)
mixed_p_lme4$r_low_alliance <- ifelse(mixed_p_lme4$alliance_Q=="Bottom_Half",
                                  mixed_p_lme4$r, 0)

# Build mixed effect linear models with:

# Random intercept per each patient

lm1_0 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (1 | ID), data=mixed_p_lme4) 

# Random slope per each patient

lm1_1 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (0 + r_high_alliance + r_low_alliance | ID), data=mixed_p_lme4) 

# Random intercept and slope per each patient

lm1_2 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (1 + r_high_alliance + r_low_alliance | ID), data=mixed_p_lme4) 

# Figure 3: plot y = changes in repayment fraction ~ x = social reciprocity values across high vs. low alliance groups.

Fig3 <-ggplot(mixed_p_lme4, aes(x=r, y=d_rf, fill=alliance_Q))+
  geom_point(aes(fill=alliance_Q))+
  facet_wrap(~alliance_Q)+
  geom_smooth(method="lm", color="black", alpha=0.5)+
  scale_fill_manual(name="Alliance Quality", 
                    values=c("Red", "Green"), 
                    labels=c("Bottom 50%", "Top 50%"))+
  xlim(-5,5)+
  labs(x="Therapist's Reciprocity", y="Fractional Change in Repayment")

ggsave("Fig3.png", height=4, width=10)

# For supplementary table 4:

# Compare thee different models

anova(lm1_0, lm1_1, lm1_2)  # winner of the three
 
# Get details of the parameter estimates from the winning model 

summary(lm1_0)
confint(lm1_0) #CI for beta regressor

plot(lm1_0) # Normality check after lmm made

# Controlling for avoidance or anxiety, does alliance still hold?

lm1_3 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + Anxiety + Avoidance + (1 | ID), data=mixed_p_lme4)

anova(lm1_0, lm1_1, lm1_2, lm1_3) #lm1_0 is still winning
summary(lm1_3)
confint(lm1_3) #CI for beta regressor

```
