---
title: "Trust_Game_in_Psychotherapy"
author: "Jihan Ryu"
date: "2024-03-12"
output: html_document
---

# 1. Load packages

```{r}
knitr::opts_chunk$set(echo = FALSE)
install.packages(c("ggplot2","dplyr","tidyr","ggcorrplot","ggpubr","statsr","lme4","lmerTest"))
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggcorrplot)
library(ggpubr)
library(statsr)
library(lme4)
library(lmerTest)
```

# 2. Upload datasets in correct format and perform descriptive stats

```{r}
# Load datasets

data <- read.csv("~/Data/Trust_Game/Github_Trust_Game/Data/TG_Raw_76.csv")

# Factor data types into correct format

factors<-c("Role","Dyad","Modality","Diagnosis","ID","Sex","Inherited","T_Experience","Medication")
integers<-c("Perception_Mismatch","Total_Alliance","Visits","Age","Task","Goal","Bond","Indiv_Total","Intrusiveness","Previous_Alliance","Lifetime_Therapist_Number","Closeness","Dependence","Anxiety","Avoidance","RFQc","RFQu","Investor_Total","Trustee_Total","Scores_Total","rt1","rt2","rt3","rt4","rt5","rt6","rt7","rt8","rt9","rt10")
numerics<-c("r3","r4","r5","r6","r7","r8","r9","r10","d_rf3","d_rf4","d_rf5","d_rf6","d_rf7","d_rf8","d_rf9","d_rf10","rf1","rf2","rf3","rf4","rf5","rf6","rf7","rf8","rf9","rf10")
data[,factors]<-lapply(data[,factors], factor)
data[,integers]<-lapply(data[,integers], as.integer)
data[,numerics]<-lapply(data[,numerics], as.numeric)
data$T_Experience <- factor(data$T_Experience, levels=c("<10","11+","100+"))

# Change missing, N/A, or #VALUE! data values into NA

data[data == "N/A" | data == "#VALUE!" | data == ""] <- NA

# For Supplementary Figure 1: create correlation matrix among numerically valued variables

temp_corr<-data[c(38:49,51:58,60:74),] #Subject 26 and 44 (patients, i.e. row 50 and 59) data are missing
temp_corr<-temp_corr %>% 
  select_if(is.numeric)
temp_corr<-temp_corr[,1:15] # select only fully available clinical/demographic variables
temp_corr$Indiv_Total_Counterpart = temp_corr$Total_Alliance - temp_corr$Indiv_Total # mutate a new variable called Indiv_Total_Counterpart, i.e. therapeutic alliance rated by their therapy partner (patient or therapist) in the pair

corr.vars <- cor(temp_corr, method="pearson")
p.mat <- cor_pmat(temp_corr)

corr_plot<-ggcorrplot(corr.vars, 
           hc.order = TRUE,
           type = "lower",
           p.mat = p.mat)

ggsave("FigS1.jpeg")
```

# 3. Mutate and extract average repayment fraction and reaction time parameters from trust game

```{r}
# Mutate and generate mean repayment values

data <- data %>% 
  rowwise() %>% 
  mutate(rf_avg = mean(as.numeric(c(rf1, rf2, rf3, rf4, rf5, rf6, rf7, rf8, rf9, rf10)), na.rm = TRUE),
         rt_avg = mean(as.numeric(c(rt1, rt2, rt3, rt4, rt5, rt6, rt7, rt8, rt9, rt10)), na.rm = TRUE))

# Split data frames into patients and therapists

data_p<-data[data$Role=="P",]
data_t<-data[data$Role=="T",]

# Summarize group mean repayment by roles (mean & SD)

rf_avg_p_mean<-colMeans(data[data$Role=="P" & is.nan(data$rf_avg)==FALSE,][,'rf_avg'])
rf_avg_p_sd<-sd(data_p$rf_avg, na.rm = TRUE)
rf_avg_t_mean<-colMeans(data[data$Role=="T",][,'rf_avg']) # Checkpoint: this value should be 0.571
rf_avg_T_sd<-sd(data_t$rf_avg, na.rm = TRUE)

# Create Figure 2

# Fig 2A: Mean pepayment frequency distribution between therapists and patients

Fig2A <- ggplot(data)+
  geom_histogram(data=data[data$Role=="P",],
                 aes(x=rf_avg), fill="Orange", alpha = 1, position="identity")+
  geom_vline(aes(xintercept = rf_avg_p_mean), colour="Orange", linetype=5)+
  geom_histogram(data=data[data$Role=="T",],
                 aes(x=rf_avg), fill="Blue", alpha = 0.8, position="identity")+
  geom_vline(aes(xintercept = rf_avg_t_mean), colour="Blue", linetype=5)+
  geom_text(data = data.frame(x = c(0.74, 0.75),
                              y = c(5, 4.5), 
                              label = c(paste("Mean P:", round(rf_avg_p_mean, 2)), 
                                        paste("Mean T:", round(rf_avg_t_mean, 2)))),
            aes(x = x, y = y, label = label),
            vjust = -0.5, size = 3.5, color = c("orange", "blue"))+ 
  labs(y="Frequency", x="Mean Repayment",
       tag="A")+
  scale_fill_manual(name="Role", 
                    values=c("Orange", "Blue"), 
                    labels=c("Patient", "Therapist"))

# Fig 2B: Correlation between mean repayment frequency and therapeutic alliance

Fig2B <- ggplot(data, aes(rf_avg, Indiv_Total, color=Role)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text(data = data.frame(x = c(0.79, 0.78),
                              y = c(84, 81), 
                              label = c("P: r=0.48, p=0.003",
                                        "T: r=0.23, p=0.15")),
            aes(x = x, y = y, label = label),
            vjust = -0.5, size = 3.5, color = c("orange", "blue"))+
  scale_color_manual(name="Role",
                     values=c("Orange", "Blue"), 
                     labels=c("Patient", "Therapist")) +
  labs(y="Alliance", x="Mean Repayment",
       tag="B")

# Combine Fig 2A, 2B into Fig 2.

Fig2 <- ggarrange(Fig2A, Fig2B, ncol=2, nrow=1, common.legend = TRUE)

ggsave("Fig2.jpeg", width=8, height=3)

# For Supplementary Figure 2: Patients' average repayment fraction is not correlated with attachment closeness subscale.

FigS2 <-ggplot(data_p, aes(rf_avg, Closeness))  +
  geom_point() +
  geom_smooth(method=lm)+
  scale_color_manual(values=c("Orange"), 
                     labels=c("Patient")) +
  labs(y="Attachment Closeness", x="Mean Repayment")

ggsave("FigS2.jpeg", width=8, height=3)

cor.test(data_p$rf_avg, data_p$Closeness, method=c("pearson"))

# For Supplementary Table 2: Create a separate data frame for reciprocity information received by participants (note: reciprocity value for r1 and r2 cannot be calculated as it requires player behavior information in the previous two rounds)

investment<-data[,c("Role","ID","r3","r4","r5","r6","r7","r8","r9","r10","I1","I2","I3","I4","I5","I6","I7","I8","I9","I10")]

# Count the number of positive & negative reciprocity rounds received by looping through values in r3-r10 

# Set counter for the number of positive & negative reciprocity rounds to zero

investment$pos = 0
investment$neg = 0

# Loop

for (j in 1:nrow(investment)){
    for (k in 3:10){
      if (is.na(investment[j,k])) {investment[j,k]=FALSE} #make sure all arguments are operating under TRUE/FALSE enabled environment
       else{
        if (investment[j,k]>0){
        investment[j,]$pos = investment[j,]$pos + 1
        }
        if (investment[j,k]<0){
        investment[j,]$neg = investment[j,]$neg + 1
        }
      }   
    }
}

# Calculate mean of raw investment received by participant throughout the 10 rounds

investment<-investment %>% 
  rowwise() %>% 
  mutate(i_avg = mean(c(I1, I2, I3, I4, I5, I6, I7, I8, I9, I10), na.rm=TRUE))

# Run inference test to see if reciprocity received was signifiantly different between patient vs. therapist groups

inference(y=neg, x=Role, data=investment, statistic="mean", type="ci", conf_level=0.95, method="theoretical", alternative="two.sided")
```

# 4. Transform data frames to perform statistical analysis among reciporicty, changes in repayment, repayment fraction variables per individual

```{r}
# First pivot, r3-10, drf3-10, and rt3-10 per ID along the rows

r<-data %>% 
  pivot_longer(
    cols=c("r3", "r4", "r5", "r6", "r7", "r8", "r9", "r10"),
    names_to="round",
    values_to="r",
    values_drop_na=FALSE
  )
d_rf<-data %>% 
  pivot_longer(
    cols=c("d_rf3", "d_rf4", "d_rf5", "d_rf6", "d_rf7", "d_rf8", "d_rf9", "d_rf10"),
    names_to="round",
    values_to="d_rf",
    values_drop_na=FALSE
  )
rf<-data %>% 
  pivot_longer(
    cols=c("rf3", "rf4", "rf5", "rf6", "rf7", "rf8", "rf9", "rf10"),
    names_to="round",
    values_to="rf",
    values_drop_na=FALSE
  )

# Combine pivoted data frames into a new data frame, i.e. 'mixed'.

mixed <- cbind(r[,c(1:26,45:47,69:72)],
               d_rf[,72],
               rf[,72]) # check if selected columns refer to correct variables

# Check distribution of data visually before running further stats

ggplot(mixed, aes(x=r, color=ID))+
  geom_histogram(binwidth=0.1)
ggplot(mixed, aes(x=d_rf, color=ID))+
  geom_histogram(binwidth=0.1)
ggplot(mixed, aes(x=rf, color=ID))+
  geom_histogram(binwidth=0.05)
```

# 5. Compare mean fractional change in repayment between benevolent rounds (reciprocity >0) and malevolent rounds (reciprocity <0)

```{r}
# Binarize reciprocity into negative (malevolence) vs positive (benevolence)

mixed$intention<-ifelse(mixed$r>0,"Benevolence",
                          ifelse(mixed$r<0, "Malevolence",NA))

# Delete rows with NA in either r or d_rf column 

mixed_filtered <- mixed %>% 
  filter(!is.na(r) & !is.na(d_rf)) # check if 75 rounds are lost due to NA.

mixed_filtered$intention <- as.factor(mixed_filtered$intention)

summary(mixed_filtered$intention) #26 neutral (r=0) rounds. 

# Figure 3: Plot group average of repayment fraction across reciprocity condtions

d_rf_intention <- mixed[is.na(mixed$intention)==FALSE,] %>% 
  group_by(intention, Role) %>% 
  summarise(sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n())%>%
  mutate(se=sd/sqrt(n))

Fig3 <- ggplot(d_rf_intention, aes(x=intention, y=mean_d_rf, fill=Role))+
  geom_bar(stat="identity", width=0.2, position=position_dodge(width=0.2)) +
  geom_errorbar(
    aes(x=intention, ymin=mean_d_rf-se, ymax=mean_d_rf+se, width=0.04), position=position_dodge(width=0.2))+
  labs(x="Therapy partner's intention", y="Mean Fractional Change in Repayment")+
  scale_x_discrete(labels = c("Positive", "Negative"))+
  scale_fill_manual(name="Role", 
                    values=c("Orange", "Blue"), 
                    labels=c("Patient", "Therapist"))

ggsave("Fig3.jpeg", width=8, height=4) 

# Divide data frames per their roles (patients vs therapists)

mixed_t <- mixed_filtered %>% 
  filter(Role == "T")
mixed_p <- mixed_filtered %>% 
  filter(Role == "P")

# Before testing significant differences between binary conditions, remove neutral rounds 

mixed_p_without_neutral <-mixed_p[is.na(mixed_p$intention)==FALSE,]
mixed_t_without_neutral <-mixed_t[is.na(mixed_t$intention)==FALSE,]

mixed_p_without_neutral %>% 
  filter(intention=="Benevolence") -> tbv
mixed_p_without_neutral %>% 
  filter(intention!="Benevolence") -> tml

var.test(tbv$d_rf, tml$d_rf) # p>0.05 in variance test indicates homoscedasticity 

# Run t.test between two conditions separately for both roles

t.test(d_rf~intention, paired=FALSE, vars.equal=TRUE,
       data=mixed_p_without_neutral) # Note: observations are not exactly matched or paired, i.e. some patients had more malevolent rounds than benevolent rounds, error in complete.cases(x,y): not all arguments have the same length. only patient data show sig differences between conditions

# Run t.test between two roles within same condition

mixed_filtered[mixed_filtered$intention=="Benevolence",] %>% 
  filter(Role=="T")->tT
mixed_filtered[mixed_filtered$intention=="Benevolence",] %>% 
  filter(Role!="T")->tP

var.test(tT$d_rf, tP$d_rf) # p>0.05 homoscedacticity for positive (benevolence), but p=0.03 for malevolence

t.test(d_rf~Role, paired=FALSE, vars.equal=TRUE,
       data=mixed_filtered[mixed_filtered$intention=="Benevolence",]) 
```

# 6. Test the effect of therapeutic alliance on repyament fraction and changes in repayment fraction for patients

```{r}
# Identify median alliance in patient group to binarize alliance variables into high vs. low alliance

test_mixed_p<-mixed_p %>% 
  distinct(ID, .keep_all = TRUE) # trim down the data frame to isolate clinical variable that is unique to each participant

summary(test_mixed_p$Indiv_Total) # median in alliance among patients should be 74. 

# Binarize alliance, attachment avoidance, attachment anxiety ratings into "Top Half" or "Top Bottom" group

mixed_p$alliance_Q<-ifelse(mixed_p$Indiv_Total>74,"Top_Half",
                               "Bottom_Half")
mixed_p$avoidance_Q<-ifelse(mixed_p$Avoidance>28,"Top_Half",
                               "Bottom_Half")
mixed_p$anxiety_Q<-ifelse(mixed_p$Anxiety>22,"Top_Half",
                               "Bottom_Half")
 
# Put binarized clinical variables into factor data type

factors_2 <- c("alliance_Q", "avoidance_Q", "anxiety_Q")

mixed_p[,factors_2]<-lapply(mixed_p[,factors_2], factor)

# Run 2 (high vs. low alliance) x 2 (positive vs. negative reciprocity) ANOVA test to examine effect of alliance on changes in repayment fraction for patients

aov_d_rf_alliance <- aov(d_rf ~ intention * alliance_Q, data=mixed_p[is.na(mixed_p$intention)==FALSE,])

summary(aov_d_rf_alliance) # not significant

# Run two extra 2 x 2 ANOVA to test impact of attachment anxiety, and by attachment avoidance on changes in repayment fraction.

aov_d_rf_anxiety <- aov(d_rf ~ intention * anxiety_Q, data=mixed_p[is.na(mixed_p$intention)==FALSE,])

summary(aov_d_rf_anxiety) 

aov_d_rf_avoidance <- aov(d_rf ~ intention * avoidance_Q, data=mixed_p[is.na(mixed_p$intention)==FALSE,])

summary(aov_d_rf_avoidance) 

# For Supplementary Table 3: Summarize the mean of changes in repayment fraction across reciprocity and alliance conditions 

d_rf_by_alliance_Q <- mixed_p[is.na(mixed_p$intention)==FALSE,] %>% 
  group_by(alliance_Q, intention) %>% 
  summarise(sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n())%>%
  mutate(se=sd/sqrt(n))

# Run correlation test between mean fractional change in repayment and individual alliance ratings within each reciprocity condition (malevolent vs. benevolent)

d_rf_by_intention_p <- mixed_p[is.na(mixed_p$intention)==FALSE,] %>%
  group_by(intention, ID) %>% 
  summarise(ID=ID, sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n(), Indiv_Total=Indiv_Total)%>%
  mutate(se=sd/sqrt(n)) %>% 
  distinct(ID, .keep_all = TRUE)

d_rf_by_anxiety_p <- mixed_p[is.na(mixed_p$intention)==FALSE,] %>%
  group_by(intention, ID) %>% 
  summarise(ID=ID, sd=sd(d_rf, na.rm=TRUE), mean_d_rf=mean(d_rf, na.rm=TRUE), n=n(), Anxiety=Anxiety)%>%
  mutate(se=sd/sqrt(n)) %>% 
  distinct(ID, .keep_all = TRUE)

attach(d_rf_by_intention_p[d_rf_by_intention_p$intention=="Benevolence",])
cor.test(mean_d_rf, Indiv_Total, alternative="two.sided", method="pearson") # not significant for linear relationship in either benevolent or malevolent condition

attach(d_rf_by_anxiety_p[d_rf_by_anxiety_p$intention=="Malevolence",])
cor.test(mean_d_rf, Anxiety, alternative="two.sided", method="pearson") # How about anxiety group since its F score was high in 2 x 2 ANOVA? Not sig in benevolent conditions, also not sig correlations shown in malevolent conditions; but trend is that anxiety increases decrease in repayment in malevolent conditions.
```

# 7. Build linear mixed effect models to predict changes in repayment fraction in patients as a function of social reprocity (as continuous variable, not categorical) and clinical variables (e.g. attachment style)

```{r}
# Remove outliers of extreme malevolence or extreme benevolence, defined by r value cut off -5 and 5, and create a separate data frame for LMM modeling.

mixed_p_lme4<-mixed_p[mixed_p$r > -5 & mixed_p$r < 5,] # Check if there are 237 observation points. 

# Create columns for reciprocity received by high alliance patient group and by low alliance patient group separately (i.e. r_high_alliance, r_low_alliance) and store them as separate predictors (regressors) 

mixed_p_lme4$r_high_alliance <- ifelse(mixed_p_lme4$alliance_Q=="Top_Half",
                                  mixed_p_lme4$r, 0)
mixed_p_lme4$r_low_alliance <- ifelse(mixed_p_lme4$alliance_Q=="Bottom_Half",
                                  mixed_p_lme4$r, 0)

# Build mixed effect linear models with:

## Random intercept per each patient

lm1_0 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (1 | ID), data=mixed_p_lme4) 

## Random slope per each patient

lm1_1 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (0 + r_high_alliance + r_low_alliance | ID), data=mixed_p_lme4) 

## Random intercept and slope per each patient

lm1_2 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + (1 + r_high_alliance + r_low_alliance | ID), data=mixed_p_lme4) 

# Figure 4: plot y = changes in repayment fraction ~ x = social reciprocity values across high vs. low alliance groups.

Fig4 <- ggplot(mixed_p_lme4, aes(x=r, y=d_rf, fill=alliance_Q))+
    geom_rect(aes(xmin=-5, xmax=0, ymin=-1, ymax=1), fill = "pink", alpha = 0.01) +
  geom_rect(aes(xmin=0, xmax=5, ymin=-1, ymax=1), fill = "skyblue", alpha = 0.01) +
  geom_point(aes(fill=alliance_Q))+
  facet_wrap(~alliance_Q, labeller = labeller(alliance_Q = c("Bottom_Half" = "Lower half alliance group", "Top_Half" = "Upper half alliance group")))+
  geom_smooth(method="lm", color="black", alpha=0.5)+
  xlim(-5,5)+
  labs(x="Simulated therapist's reciprocity", y="Fractional Change in Repayment")+
  theme(legend.position = "none") 

ggsave("Fig4.jpeg", width=8, height=4)

# For Supplementary Table 4: compare the the three models to find the best fitting one

anova(lm1_0, lm1_1, lm1_2)  # winner of the three is lm1_0 (random intercept only)
 
# Get details of the parameter estimates from the winning model 

summary(lm1_0)
confint(lm1_0) # CI for beta regressor
plot(lm1_0) # Normality/residuals check

# Controlling for avoidance or anxiety, does alliance still hold?

lm1_3 <- lmer(d_rf ~ 1 + r_high_alliance + r_low_alliance + Anxiety + Avoidance + (1 | ID), data=mixed_p_lme4)

anova(lm1_0, lm1_1, lm1_2, lm1_3) #lm1_0 is still winning
summary(lm1_3)
confint(lm1_3) #CI for beta regressor

# End
```
```{r}
# Converting plot figures into .jpeg of 300 dpi or higher 

library(magick)

img <- image_read("Fig2.jpeg")
image_info(img)

ggsave("Fig2.jpeg", Fig2, width = 8, height = 3, dpi = 300)
ggsave("Fig3.png", Fig3, width = 8, height = 4, dpi = 300)
ggsave("Fig4.png", Fig4, width = 8, height = 4, dpi = 300)


```

